A Kochen-Specker system has at least 21 vectors\\ {\small (extended abstract)}

### 2014-09-12 paper/main.tex
paper: suggesties van Sander

\begin{dfn}
Given a finite subset~$S$ of the projective plane
(or equivalently, a finite subset of the northern
hemisphere without equator).
Define its \keyword{orthogonality graph}~$G(S)$ as follows.
The vertices are the points of~$S$.
Two vertices are joined by an edge, if their corresponding points
are orthogonal.
\end{dfn}
\begin{dfn}
A graph~$G$ is called~\keyword{embeddable},
if it occurs as a subgraph of an orthogonality graph;
that is: if there is a finite subset~$S$ of the projective plane,
such that~$G \leq G(S)$.
\end{dfn}
% TODO add remark on subgraph definition?
\begin{dfn}
A graph is~\keyword{010-colorable}
if there is a~$\{0,1\}$-coloring, such that
\begin{enumerate}
\item
for each triangle there is exactly one vertex that is colored~$1$ and
\item
adjacent vertices are not both colored~$1$.
\end{enumerate}
\end{dfn}
\begin{dfn}
A \keyword{Kochen-Specker graph}
is a embeddable graph that is not 010-colorable.
\end{dfn}
It is an easy, but important, consequence of the definitions that:
\begin{fact}
    A finite subset~$S$ of the projective plane
    is a Kochen-Specker system,
    if and only if its orthogonality graph~$G(S)$
    is Kochen-Specker.
\end{fact}
To prove there is no Kochen-Specker system on~$17$ points,
it would be sufficient to enumerate all graphs on~$17$ vertices
and check these are not 010-colorable or not embeddable.
However, this is infeasible as there are
already~${\sim}10^{26}$ non-isomorphic
graphs on~$17$ points.\cite{oeisA000088}
Luckily, we can restrict ourselves to certain classes of graphs.
\begin{prop}[\cite{aow11}]
    An embeddable graph is squarefree.
    That is: it does not contain the square as a subgraph.%
    \footnote{Some authors call a graph squarefree if it does not
        contain the square as induced subgraph.
        For them the complete graph on four vertices is squarefree.
        We follow~Weisstein\cite{sf-weisstein} and~Sloane\cite{sf-sloane} and
        call a graph squarefree if it does not
        contain the square as subgraph.
        For us the complete graph on four vertices is not squarefree.}
\end{prop}
\begin{proof}
    Given two non antipodal points~$v\neq w$.
    See Figure~\ref{fig:crossproduct}.
    Consider the points orthogonal to~$v$.
    This is a great circle.
    The points orthogonal to~$w$ is a different great circle.
    They intersect in precisely two antipodal points.
    Hence, if~$c$ and~$d$ are both orthogonal to~$v$ and~$w$,
    then~$c$ and~$d$ are equivalent.
    Therefore, an embeddable graph cannot contain a square.
\end{proof}
The squarefreeness is a considerable restriction.  There are
only~${\sim}10^{10}$ non-isomorphic squarefree graphs on~$17$
vertices.\cite{sf-sloane}
We can restrict ourselves to connected graphs.
\begin{prop}[\cite{aow11}]\label{prop:ks-conn}
    A minimal Kochen-Specker graph is connected.
\end{prop}
\begin{proof}
    Suppose~$G$ is a non-connected Kochen-Specker graph.
    Then one of its components is not 010-colorable.
    As a subgraph of an embeddable graph, is embeddable,
    this component is embeddable as well.
    Hence it is a smaller connected Kochen-Specker graph.
\end{proof}
The gain, however, is small.
There are only~${\sim}10^9$ non-isomorphic squarefree graphs on~$17$
vertices that are not connected.
In our computations, checking for connectedness
required more time than would be gained by reducing the number of graphs.

@@

\section{An improved lower bound}
\label{sec:ilb}
Continuing the effort of Arends, Ouaknine and Wambler,
we consider another restriction.
\begin{prop}
    A minimal Kochen-Specker graph has minimal vertex-order three.
\end{prop}
\begin{proof}
    Given a Kochen-Specker graph~$G$.
    Suppose~$v$ is a vertex with order less than or equal~$2$.
    Let~$G'$ be~$G$ with~$v$ removed.
    Clearly~$G'$ is embeddable.
    Suppose~$G'$ is 010-colorable.
    Then we can extend the coloring to a coloring of~$G$ as follows.
    If~$v$ is adjacent to only one or no vertex,
    then we can color~$v$ with~$0$.
    Suppose~$v$ is adjacent to two vertices, say~$w$ and~$w'$.
    If one of~$w$ or~$w'$ is colored~$1$, we can color~$v$ with~$0$.
    If both~$w$ and~$w'$ are colored~$0$, we can color~$v$ with~$1$.
    This would imply~$G$ is 010-colorable, quod non.
    Therefore~$G'$ is a smaller
    Kochen-Specker graph, which contradicts minimality.
\end{proof}
There are only~${\sim}10^7$
squarefree non-isomorphic graphs on 17 vertices with minimal vertex order 3.
Even though Arends, Ouaknine and Wampler
note this restriction once,
surprisingly, they did not restrict their graph enumeration
to graphs with minimal vertex order 3.

@@

\begin{dfn}
   Given a graph~$G$ together with selected vertices~$v_1, \ldots, v_n \in G$.
   Let~$C(G)$ denote the set of $010$-colorings of~$G$.
   The \keyword{type of~$(v_1,\ldots,v_n)$ (in~$G$)}
   is the set of all possible ways it can be colored;
   that is: $t_G(v_1, \ldots, v_n) := \{ (c(v_1), \ldots, c(v_n));
                \ c \in C(G) \}.$ 
    A type of~$n$ vertices is
    called an~\keyword{$n$-type}.
\end{dfn}

Note that~$v$ has the~$1$-type~$(1)$ in~$G$ if and only if
it has fixed color~$1$ in~$G$.
Just like points with fixed color,
we are interested in types,
because they do not occur in small graphs.

\begin{comp}\label{comp:types}
    We have enumerated all embeddable graphs of less than~$17$ vertices
    and determined a lower bound at which a particular $1$- or $2$-type occurs.
    \begin{center}
    \begin{tabular}{ll}
        $1/2$-type & $\#G$ \\ \hline
        $\{(0,0), (1,0), (0,1)\}$ &
            non-trivially $\geq 10$\\ 
        $\{(0,0), (1,0), (1,1)\}$ & $\geq 10$ \\
        $\{(0,0), (0,1), (1,1)\}$ & $\geq 10$ \\
        $\{(0,0), (0,1))\}$ & $\geq 15$ \\
        $\{(0,0), (1,0))\}$ & $\geq 15$ \\
        $\{(0)\}$ & $\geq 15$ \\
        $\{(0,1),(1,0)\}$ & $\geq 16$ \\
        other & $\geq 17$
    \end{tabular}
    \end{center}
    The type~$\{(0,0),(1,0),(0,1)\}$ occurs in the embeddable two-vertex
    graph
    \begin{tikzpicture}[thick,scale=1.4,
        p/.style={circle, draw, fill=white,
                        inner sep=0pt, minimum width=5pt}]
            \draw
            (0,0) node[p]{} -- (0.3,0) node[p]{} ;
    \end{tikzpicture}.
    Because the two vertices are adjacent, this occurance of
    the type is called trivial.
\end{comp}

@@

\section{Embeddability}
Our computation has yielded a few hundred non-010-colorable graphs.
If we show one of them is embeddable, we have found a new KS system.
If we demonstrate all of them are not embeddable, we have
proven a lower bound on the size of a minimal KS system.

@@

We propose a new method,
which for all graphs we considered,
could decide
whether they were embeddable or not.
First we give a pen-and-paper example.
\begin{prop}\label{prop:unemb-10-2}
The graph in Figure~\ref{fig:unemb-10-2}
is not embeddable.
\end{prop}
\begin{proof}
Suppose it is embeddable.
Consider~$p_1$.
It is orthogonal to both~$a$ and~$v$.
$a$ and~$v$ are not collinear,
hence~$p_1$ must be collinear to~$v \times a$,
the cross-product of~$v$ and~$a$.
Similarly, $p_2$ is collinear to~$v \times p_1 = v \times (v \times a)$.
Continuing in this fashion,
we see that
\begin{equation}\label{eq:ue1}
    a \text{ is collinear to }
    x \times (x \times( w \times (w\times (v \times (v \times a))))).
\end{equation}
Now, we may assume that~$z=(0,0,1)$ and $x=(1,0,0)$.
Thus: $v=(v_1,v_2,0)$;
$w = (w_1,w_2,0)$
and $a = (0, a_2,a_3)$ for some~$-1 \leq v_1,v_2,w_1,w_2,a_2,a_3 \leq 1$,
with~$v_1^2+v_2^2 = 1$; $w_1^2+w_2^2=1$ and~$a_2^2 + a_3^2=1$.
Now, \eqref{eq:ue1} becomes:
\begin{equation*}
\begin{pmatrix}
0\\
a_2\\
a_3\\
\end{pmatrix}
\text{ is collinear to }
\begin{pmatrix}
0\\
-a_2 v_1w_2 (v_1w_1 + v_2w_2) \\
-a_3 (v_1^2 w_1^2 + v_1^2 w_2^2 + v_2^2w_1^2 + v_2^2w_2^2)\\
\end{pmatrix}.
\end{equation*}
And therefore
\begin{align*}
v_1w_2 \left<v,w\right> & =
v_1w_2 (v_1w_1 + v_2w_2) \\
& = v_1^2 w_1^2 + v_1^2 w_2^2 + v_2^2w_1^2 + v_2^2w_2^2 \\
& = (v_1^2 + v_2^2)w_1^2 + (v_1^2 + v_2^2)w_2^2 \\
& = w_1^2 + w_2^2 = 1.
\end{align*}
Since~$v$ and~$w$ are not collinear,
we have by Cauchy-Schwarz $|\left<v,w\right>| < 1$.
Recall~$|v_1|, |w_2| \leq 1$.
Thus: $|v_1w_2\left<v,w\right>| < 1$.
Contradiction.
\end{proof}

@@

The requirements can be mechanically converted
to a formal sentence
in the language of the real numbers.
This sentence is true if and only if the graph is embeddable.
Famously, Tarski proved\cite{tarski}
that such sentences are decidable.
His decision procedure has an impractical complexity.
However, its practical value has been improved
by, for instance, the method of cylindrical algebraic decomposition\cite{qecad}.
We have used the redlog\cite{redlog} package of the reduce algebra
system, which implements a variant of Tarski's quantifier elimination.
In Appendix~\ref{adx:unemb-10-2} the Reader can find
the reduce script generated mechanically for the graph
in Figure~\ref{fig:unemb-10-2}.

@@

In this way, there were still a few (010-colorable) graphs of which
we could not decide embeddability.
By hand, we determined these graphs to be embeddable.
We adapted the algorithm, as to guess
for some assignments the position of one of the vectors.
If the corresponding
sentence turns out false, we know nothing.  However,
if the sentence is true, we know the graph is embeddable.

With this method, we have decided in an a day the embeddability
of every squarefree graph with minimal vertex order three of 14
vertices of less except for one.\footnote{
A list of all squarefree graphs with minimal vertex order three of
14 vertices of less together with their embeddability can
be found here:
\url{http://kochen-specker.info/smallGraphs/}
}
In particular:
\begin{comp}\label{comp:unemb20}
    Every squarefree graph of minimal vertex order three
    that is not 010-colorable
    of order less than or equal to 20
    contains, as a subgraph, one of the following three graphs:
    \begin{center}
    \begin{tikzpicture}[thick,scale=1.4,
        p/.style={circle, draw, fill=white,
                        inner sep=0pt, minimum width=5pt}]
            \draw
                \foreach \x in {90,210,330} {
                    (\x:0.7) node[p]{} -- (0:0) node[p]{}
                    (\x-30:1.3) node[p]{} -- (\x:0.7)
                    (\x+30:1.3) node[p]{} -- (\x:0.7)
                    (\x-30:1.3) -- (\x+30:1.3) 
                    (\x+30:1.3) -- (\x-30+120:1.3) 
                }
            ;
        \end{tikzpicture}
        \qquad
    \begin{tikzpicture}[thick,scale=1.4,
        p/.style={circle, draw, fill=white,
                        inner sep=0pt, minimum width=5pt}]
            \draw
                \foreach \x in {0,180} {
                    (\x+15:1.3) node[p]{} -- (\x+90:0.9) node[p]{}
                    (\x+15:1.3) -- (\x+45:0.5) node[p]{}
                    (\x+45:0.5) -- (\x+90:0.9)

@@

We are grateful to prof.~McKay for discussing
the feasibility of certain graph restrictions and
to Judith van Stegeren\footnote{\url{http://jd7h.com/}}
for drawing the first two figures in this paper.

\clearpage
\appendix
\section{Example reduce script}\label{adx:unemb-10-2}
The following is (a part of) the reduce script mechanically generated
to prove the non-embeddability of the graph of Figure~\ref{fig:unemb-10-2}.
The algorithm choose a different assignment, than we did in the
proof of Proposition~\ref{prop:unemb-10-2}.
As free points it picked, in order,~$z$, $x$, $v$ and~$p_3$.
The point~$w$ is assigned~$p_3 \times z$.
\begin{verbatim}
load_package redlog;
rlset R;

procedure d(x,y);
    (first x) * (first y) +
    (second x) * (second y) +
    (third x) * (third y);

procedure k(x,y);
    {(second x)*(third y) - (third x)*(second y),
     (third x)*(first y) - (first x)*(third y),
     (first x)*(second y) - (second x)*(first y)};

v0c1 := 1; v0c2 := 0; v0c3 := 0;
v1c1 := 0; v1c2 := 1; v1c3 := 0;

v0 := {v0c1, v0c2, v0c3}; 
v1 := {v1c1, v1c2, v1c3}; 
v2 := {v2c1, v2c2, v2c3}; 
v3 := {v3c1, v3c2, v3c3}; 
v2c1 := 0;
neq0 := k(v0,k(v3,v1)); 
\end{verbatim}
\begin{center} \emph{(snip)} \end{center}
\begin{verbatim}
neq29 := k(k(k(k(v3,v1),v1),v2),k(k(v3,v0),v3)); 
phi := 
       (first neq0 neq 0 or
        second neq0 neq 0 or
        third neq0 neq 0) and 
\end{verbatim}
\begin{center} \emph{(snip)} \end{center}
\begin{verbatim}
       (first neq29 neq 0 or
        second neq29 neq 0 or
        third neq29 neq 0) and 
       d(v2,v0) = 0 and 
       d(k(k(v3,v0),v3),k(k(k(k(v3,v1),v1),v2),v2)) = 0 and 
        true;
rlqe ex(v3c3,
     ex(v3c2,
     ex(v3c1,
     ex(v2c3,
     ex(v2c2,phi)))));
\end{verbatim}

>>>>

\begin{dfn}
Given a finite subset~$S$ of the projective plane
(or equivalently, a finite subset of the northern
hemisphere without equator\footnote{%
    A subset of the projective plane can be identified with
    a subset of the closed northern hemisphere.
    For a finite subset we can always rotate in such a way
    that no points lie on the equator.}).
Define its \keyword{orthogonality graph}~$G(S)$ as follows.
The vertices are the points of~$S$.
Two vertices are joined by an edge, if their corresponding points
are orthogonal.
\end{dfn}
\begin{dfn}
A graph~$G$ is called~\keyword{embeddable},
if it occurs as a subgraph of an orthogonality graph;
that is: if there is a finite subset~$S$ of the projective plane,
such that~$G \leq G(S)$.
\end{dfn}
% TODO add remark on subgraph definition?
\begin{dfn}
A graph is called~\keyword{010-colorable}
if there is a~$\{0,1\}$-coloring of the vertices,
such that
\begin{enumerate}
\item
for each triangle there is exactly one vertex that is colored~$1$ and
\item
adjacent vertices are not both colored~$1$.
\end{enumerate}
\end{dfn}
\begin{dfn}
A \keyword{Kochen-Specker graph}
is a embeddable graph that is not 010-colorable.
\end{dfn}
It is an easy, but important, consequence of the definitions that:
\begin{fact}
    A finite subset~$S$ of the projective plane
    is a Kochen-Specker system,
    if and only if its orthogonality graph~$G(S)$
    is Kochen-Specker.
\end{fact}
To prove there is no Kochen-Specker system on~$17$ points,
it would be sufficient to enumerate all graphs on~$17$ vertices
and check these are not 010-colorable or not embeddable.
However, this is infeasible as there are
already~${\sim}10^{26}$ non-isomorphic
graphs on~$17$ points.\cite{oeisA000088}
Luckily, we can restrict ourselves to certain classes of graphs.
\begin{prop}[\cite{aow11}]
    An embeddable graph is squarefree.
    That is: it does not contain the square as a subgraph.%
    \footnote{Some authors call a graph squarefree if it does not
        contain the square as induced subgraph.
        For them the complete graph on four vertices is squarefree.
        We follow~Weisstein\cite{sf-weisstein} and~Sloane\cite{sf-sloane} and
        call a graph squarefree if it does not
        contain the square as subgraph.
        For us the complete graph on four vertices is not squarefree.}
\end{prop}
\begin{proof}
    Given two non antipodal points~$v\neq w$.
    See Figure~\ref{fig:crossproduct}.
    Consider the points orthogonal to~$v$.
    This is a great circle.
    The points orthogonal to~$w$ is a different great circle.
    They intersect in precisely two antipodal points.
    Hence, if~$c$ and~$d$ are both orthogonal to~$v$ and~$w$,
    then~$c$ and~$d$ are equivalent.
    Therefore, an embeddable graph cannot contain a square.
\end{proof}
The squarefreeness is a considerable restriction.  There are
only~${\sim}10^{10}$ non-isomorphic squarefree graphs on~$17$
vertices.\cite{sf-sloane}
Next, we show we can restrict ourselves to connected graphs.
\begin{prop}[\cite{aow11}]\label{prop:ks-conn}
    A minimal Kochen-Specker graph is connected.
\end{prop}
\begin{proof}
    Suppose~$G$ is a non-connected Kochen-Specker graph.
    Then one of its components is not 010-colorable.
    As a subgraph of an embeddable graph, is embeddable,
    this component is embeddable as well.
    Hence it is a smaller connected Kochen-Specker graph.
\end{proof}
The gain, however, is small.
There are only~${\sim}10^9$ non-isomorphic squarefree graphs on~$17$
vertices that are not connected.
In our computations, checking for connectedness
required more time than would be gained by reducing the number of graphs.

@@

\section{An improved lower bound}
\label{sec:ilb}
Continuing the effort of Arends, Ouaknine and Wambler,
we consider another restriction.
\begin{prop}
    A minimal Kochen-Specker graph has minimal vertex-order three.
    That is: every vertex is adjacent to at least three other vertices.
\end{prop}
\begin{proof}
    Given a minimal Kochen-Specker graph~$G$.
    Suppose~$v$ is a vertex with order less than or equal~$2$.
    Let~$G'$ be~$G$ with~$v$ removed.
    Clearly~$G'$ is embeddable.
    Suppose~$G'$ is 010-colorable.
    Then we can extend the coloring to a coloring of~$G$ as follows.
    If~$v$ is adjacent to only one or no vertex,
    then we can color~$v$ with~$0$.
    Suppose~$v$ is adjacent to two vertices, say~$w$ and~$w'$.
    If one of~$w$ or~$w'$ is colored~$1$, we can color~$v$ with~$0$.
    If both~$w$ and~$w'$ are colored~$0$, we can color~$v$ with~$1$.
    This would imply~$G$ is 010-colorable, quod non.
    Therefore~$G'$ is a smaller
    Kochen-Specker graph, which contradicts the minimality of~$G$.
\end{proof}
There are only~${\sim}10^7$
squarefree non-isomorphic graphs on 17 vertices with minimal vertex order 3.
Even though Arends, Ouaknine and Wampler
note this restriction once,
surprisingly, they did not restrict their graph enumeration
to graphs with minimal vertex order 3.

@@

\begin{dfn}
   Given a graph~$G$ together with selected vertices~$v_1, \ldots, v_n \in G$.
   Let~$C(G)$ denote the set of $010$-colorings of~$G$.
   The \keyword{type~$t$ of~$(v_1,\ldots,v_n)$ (in~$G$)}
   is the set of all possible ways $\{v_1, \ldots, v_n\}$ can be colored;
   that is: $t = \{ (c(v_1), \ldots, c(v_n));
                \ c \in C(G) \}.$ 
    A type of~$n$ vertices is
    called an~\keyword{$n$-type}.
\end{dfn}

\begin{exa}
\begin{itemize}
\item The triangle has~$3$-type~$\{(1,0,0), (0,1,0), (0,0,1)\}$.
\item Every vertex in a Kochen-Specker graph has type~$\emptyset$.
\item
    A vertex~$v$ has the~$1$-type~$\{(1)\}$ in~$G$ if and only if
        it has fixed color~$1$ in~$G$.
\end{itemize}
\end{exa}

Just as verteces with fixed color are rare,
we are interested in types,
because most types do not occur in small graphs.

\begin{comp}\label{comp:types}
    We have enumerated all embeddable graphs of less than~$17$ vertices
    and determined a lower bound at which a particular $1$- or $2$-type occurs,
    omitting the trivial types~$\{(0),(1)\}$
    and~$\{(0,0),(0,1),(1,0),(1,1)\}$.
    \begin{center}
    \begin{tabular}{ll}
        $1/2$-type & $\#G$ \\ \hline
        $\{(0,0), (1,0), (0,1)\}$ &
            non-trivially $\geq 10$\\ 
        $\{(0,0), (1,0), (1,1)\}$ & $\geq 10$ \\
        $\{(0,0), (0,1), (1,1)\}$ & $\geq 10$ \\
        $\{(0,0), (0,1))\}$ & $\geq 15$ \\
        $\{(0,0), (1,0))\}$ & $\geq 15$ \\
        $\{(0)\}$ & $\geq 15$ \\
        $\{(0,1),(1,0)\}$ & $\geq 16$ \\
        other & $\geq 17$
    \end{tabular}
    \end{center}
    The type~$\{(0,0),(1,0),(0,1)\}$ occurs in the embeddable two-vertex
    graph
    \begin{tikzpicture}[thick,scale=1.4,
        p/.style={circle, draw, fill=white,
                        inner sep=0pt, minimum width=5pt}]
            \draw
            (0,0) node[p]{} -- (0.3,0) node[p]{} ;
    \end{tikzpicture}.
    Because the two vertices are adjacent, this occurance of
    the type is called trivial.
\end{comp}

@@

\section{Embeddability}
Our computation has yielded over nine-thousand non-010-colorable graphs.
If we show one of them is embeddable, we have found a new KS system.
If we demonstrate all of them are not embeddable, we have
proven a lower bound on the size of a minimal KS system.

@@

We propose a new method,
which for all graphs we considered,
could decide
whether they were embeddable or not.
First we give a pen-and-paper example.
\begin{prop}\label{prop:unemb-10-2}
The graph in Figure~\ref{fig:unemb-10-2}
is not embeddable.
\end{prop}
\begin{proof}
Suppose it is embeddable.
Consider~$p_1$.
It is orthogonal to both~$a$ and~$v$.
Since $a$ and~$v$ are not collinear,
$p_1$ must be collinear to~$v \times a$,
the cross-product of~$v$ and~$a$.
Similarly, $p_2$ is collinear to~$v \times p_1 = v \times (v \times a)$.
Continuing in this fashion,
we see that
\begin{equation}\label{eq:ue1}
    a \text{ is collinear to }
    x \times (x \times( w \times (w\times (v \times (v \times a))))).
\end{equation}
Now, we may assume that~$z=(0,0,1)$ and $x=(1,0,0)$.
Thus: $v=(v_1,v_2,0)$;
$w = (w_1,w_2,0)$
and $a = (0, a_2,a_3)$ for some~$-1 \leq v_1,v_2,w_1,w_2,a_2,a_3 \leq 1$,
with~$v_1^2+v_2^2 = 1$; $w_1^2+w_2^2=1$ and~$a_2^2 + a_3^2=1$.
Now, \eqref{eq:ue1} becomes:
\begin{equation*}
\begin{pmatrix}
0\\
a_2\\
a_3\\
\end{pmatrix}
\text{ is collinear to }
\begin{pmatrix}
0\\
-a_2 v_1w_2 (v_1w_1 + v_2w_2) \\
-a_3 (v_1^2 w_1^2 + v_1^2 w_2^2 + v_2^2w_1^2 + v_2^2w_2^2)\\
\end{pmatrix}.
\end{equation*}
Consequently
\begin{align*}
    v_1w_2 (v_1w_1 + v_2w_2)
& = v_1^2 w_1^2 + v_1^2 w_2^2 + v_2^2w_1^2 + v_2^2w_2^2 \\
& = (v_1^2 + v_2^2)w_1^2 + (v_1^2 + v_2^2)w_2^2 \\
& = w_1^2 + w_2^2 \\
&= 1.
\end{align*}
Since~$v$ and~$w$ are not collinear,
we have by Cauchy-Schwarz $|\left<v,w\right>| < 1$.
Now we find the contradiction:
\begin{equation*}
1 > |v_1w_2\left<v,w\right>| = |v_1w_2(v_1w_1+v_2w_2)| = 1. \qedhere
\end{equation*}
\end{proof}

@@

The requirements can be mechanically converted
to a formal sentence
in the language of the real numbers.
This sentence is true if and only if the graph is embeddable.
Famously, Tarski proved\cite{tarski}
that such sentences are decidable.
His decision procedure has an impractical complexity.
However, its practical value has been improved
by, for instance, the method of cylindrical algebraic decomposition\cite{qecad}.
We have used the redlog\cite{redlog} package of the reduce algebra
system, which implements a variant of Tarski's quantifier
elimination.\footnote{%
    The reader can find the reduce script generated mechanically
    for the graph in Figure~\ref{fig:unemb-10-2} here:
    \url{http://kochen-specker.info/smallGraphs/49743f49514769444f.html}.  }

@@

In this way, there were still a few (010-colorable) graphs of which
we could not decide embeddability.
With some guessing,
we determined embeddings for these graphs by hand.
Once we knew the troublesome graphs were embeddable,
we adapted the algorithm, as to guess
for some assignments the position of one of the vectors.
If the corresponding
sentence turns out false, we know nothing.  However,
if the sentence is true, we know the graph is embeddable.

With this method, we have decided in a day the embeddability
of every squarefree graph with minimal vertex order three of
less than 15, except for one.\footnote{
A list of all squarefree graphs with minimal vertex order three of
less than 15 vertices together with their embeddability can
be found here:
\url{http://kochen-specker.info/smallGraphs/}.
The graph for which we could not determine
embeddability can be found
here: \url{http://kochen-specker.info/smallGraphs/4d4b3f4b3f603f47414641654953625f3f.html}.
}
In particular:
\begin{comp}\label{comp:unemb20}
    Every squarefree graph of minimal vertex order three
    that is not 010-colorable
    of order less than or equal to 20
    contains, as a subgraph, one of the following three graphs:
    \begin{center}
    \begin{tikzpicture}[thick,scale=1.4,
        p/.style={circle, draw, fill=white,
                        inner sep=0pt, minimum width=5pt}]
            \draw
                \foreach \x in {90,210,330} {
                    (\x:0.7) node[p]{} -- (0:0) node[p]{}
                    (\x-30:1.3) node[p]{} -- (\x:0.7)
                    (\x+30:1.3) node[p]{} -- (\x:0.7)
                    (\x-30:1.3) -- (\x+30:1.3) 
                    (\x+30:1.3) -- (\x-30+120:1.3) 
                }
            ;
        \end{tikzpicture}
        \qquad
    \begin{tikzpicture}[thick,scale=1.4,
        p/.style={circle, draw, fill=white,
                        inner sep=0pt, minimum width=5pt}]
            \draw
                \foreach \x in {0,180} {
                    (\x+15:1.3) node[p]{} -- (\x+90:0.9) node[p]{}
                    (\x+15:1.3) -- (\x+45:0.5) node[p]{}
                    (\x+45:0.5) -- (\x+90:0.9)

@@

We are grateful to prof.~McKay for discussing
the feasibility of certain graph restrictions and
to Judith van Stegeren\footnote{\url{http://jd7h.com/}}
for drawing figures for the draft version of this paper.
